---
title: "Econ294 Final Exam Report"
author: "Naihao Guo"
date: "2016/03/16"
output: pdf_document
---

__Starting__

```{r,warning=FALSE,eval=TRUE,message=FALSE,message=FALSE}
library(dplyr)
library(ggplot2)
library(nycflights13)
my_db <- src_sqlite("db.sqlite11", create = T)
flights_sqlite <- copy_to(my_db, flights, temporary = FALSE, indexes = list(
  c("year", "month", "day"), "carrier", "tailnum"))
```

__a) weather__

```{r,warning=FALSE,eval=TRUE,message=FALSE,message=FALSE}
my_db2 <- src_sqlite("db.sqlite12", create = T)
weather_sqlite <- copy_to(my_db2, weather, temporary = FALSE, indexes = list(
  c("year", "month", "day"), "origin"))
flights_delay <- tbl(my_db, sql(
  "SELECT * FROM flights"))%>% collect()%>% mutate(canceled=is.na(arr_time))
weather <- tbl(my_db2, sql("SELECT * FROM weather"))%>% collect()
weather_delay_cancel <- inner_join(flights_delay,weather)
weather_delay_cancel <- as.data.frame(weather_delay_cancel)

weather.dep.fit <- lm(dep_delay ~ temp +  humid + wind_speed + visib,
                      data = weather_delay_cancel)
summary(weather.dep.fit)
```

From the above output, we could see that departure delay in year 2003 is effected by many factors, the significant ones with p values less than 0.05 are Temperature,Relative humidity,wind speed and visibility. So according to the output, it can be seen that 1 additional unit Temperature increase is associated with 0.02076 minutes delay in departure time,  1 additional unit increase in Relative humidity is associated with -0.13 minutes delay in departure time, 1 additional unit increase in wind speed is associated with 0.073 minutes delay in departure time, 1 addtional unit increase in visibility is associated with -3.077 minutes delay in departure time.


```{r,warning=FALSE,eval=TRUE,message=FALSE}
weather.cancel.fit <- glm(
  canceled ~ temp +  humid + wind_speed + visib,
  data = weather_delay_cancel, family = binomial(link="logit"))
summary(weather.cancel.fit)
```

Temparture and visibility significantly effect cancelations.


__b) time of day, day of week, and time of year, and any other aspect of time__


```{r,warning=FALSE,eval=TRUE,message=FALSE}
flights_delay2 <- tbl(my_db, sql("SELECT * FROM flights"))%>% mutate(canceled = is.na(arr_time))
flights_delay2 <- as.data.frame(flights_delay)
flights_delay_df <- tapply(flights_delay2$dep_delay,flights_delay2$hour, mean, na.rm=T )
flights_delay_df2 <- data.frame(hour=0:24, meandelay=flights_delay_df)
fig <- ggplot(flights_delay_df2 , aes(
  x= factor(hour),y=meandelay)) + geom_bar(stat="identity", fill ="skyblue")
fig
time.dep.fit.hour <- lm(dep_delay ~ hour,data = flights_delay)
summary(time.dep.fit.hour)
```

From the mean value departure delay and hours barplot, it can be seen hour seems to be a significant factor, and hour 3 has a most high average departure delay time.

From the output of the model, We can see that the p value of hour is very close to 0, so the hour is significant factor effect departure delay, and one additional hour is associated with 1.6594 minuses increase in departure delay.

```{r}
time.cancel.fit.hour <- glm(canceled ~ as.factor(hour), data = flights_delay, family = binomial(link ="logit"))
summary(time.cancel.fit.hour)
```

So now we can see the p value of hour, so it is not a significant factor effect cancelations.

```{r}
time.dep.fit.month <- lm(dep_delay ~ month,data = flights_delay)
summary(time.dep.fit.month)
```

We can see that the p value of month is close to 0, so the month is significant factor effect departure delay, and one additional month is associated with -0.23634 minuses increase in departure delay.

```{r}
time.cancel.fit.month <- glm(canceled ~ as.factor(month),
                             data = flights_delay, family = binomial(link ="logit"))
summary(time.cancel.fit.month)
```

So now we can see that Feb, Mar, Apr, June, July, and Dec significantly affect the cancelation with a positive influence, while Oct and Nov have negative influence on cancelation.


```{r}
flights_delay$quarter <- ifelse(flights_delay$month %in% c(1,2,3),1, ifelse(
  flights_delay$month %in% c(4,5,6),2, ifelse(flights_delay$month %in% c(7,8,9),3,4)))

time.dep.fit.quarter <- lm(dep_delay ~ quarter,data = flights_delay)
summary(time.dep.fit.quarter)
```

We can see that the p value of quarter is 2e-16 < 0.05, so the quarter is significant factor effect departure delay, and one additional quarter is associated with -0.86477  minuses increase in departure delay.

```{r}
time.cancel.fit.quarter <- glm(canceled ~ as.factor(quarter),
                               data = flights_delay, family = binomial(link = "logit"))
summary(time.cancel.fit.quarter )
```

Here we can see that the p value of quarters are small, so they are significant factors affect cancelations.


```{r}
months <- ifelse(flights_delay$month < 10, paste0("0",flights_delay$month),flights_delay$month)
days <- ifelse(flights_delay$day < 10, paste0("0",flights_delay$day),flights_delay$day)
flights_delay$dayofweek <- strftime(as.Date(paste0(flights_delay$year,"-",months,"-",days)),"%u")
flights_delay_df <- tapply(flights_delay$dep_delay,flights_delay$dayofweek, mean, na.rm=T )
flights_delay_df2 <- data.frame(dayofweek=1:7, meandelay=flights_delay_df)
fig <- ggplot(flights_delay_df2 , aes(
  x= factor(dayofweek),y=meandelay)) + geom_bar(stat="identity", fill = "skyblue")
fig
time.dep.fit.dayofweek <- lm(dep_delay ~ as.integer(dayofweek),data = flights_delay)
summary(time.dep.fit.dayofweek)
```

From the mean value departure delay and day of week barplot, it can be seen day of week seems to be a significant factor, and day of week  4 has a most high average departure delay time while day of week 6 has a most low average departure delay time.

From the model output, We can see that the p value of dayofweek is less than 0.05, so the dayofweek is a significant factor effect departure delay.

```{r}
time.cancel.fit.dayofweek <- glm(canceled ~ as.integer(dayofweek),
                                 data = flights_delay, family = binomial(link = "logit"))
summary(time.cancel.fit.dayofweek)
```

Here we can see that the p value of dayofweek is  3.39e-8 < 0.05, so it is a significant factor effect cancelations.



__c) airport destination__


```{r,warning=FALSE,eval=TRUE,message=FALSE}
flights_delay <- tbl(my_db, sql("SELECT dep_delay,arr_time, year, dest FROM flights")
                     ) %>% filter(year == 2013) %>% mutate(canceled = is.na(arr_time))
flights_delay <- as.data.frame(flights_delay)
dest.dep.fit <- lm(dep_delay ~ dest, data= flights_delay)
summary(dest.dep.fit)
```

From the plot, it can be easily seen that some destinations have large average departure delay, and there is one destination seems has the most early average departure time.

From the above model ouput,  there are some destinations would significantly effect departure delays, for example, dest ALB has p values less than 0.05, so th destination is significant factor, and the estimated value is  23.73511, so it means it is about departure delay 23.73511 minutes more than the base line destination.


```{r}
tmp <- group_by(flights, dest) %>% summarise(meandelay = mean(dep_delay,na.rm=TRUE))
tmp <- as.data.frame(tmp)
fig <- ggplot(tmp, aes(x= dest,y=meandelay)) + geom_bar(stat="identity",fill ="skyblue") + 
                                                theme(axis.text.x=element_text(angle = 90))
fig
dest.cancel.fit <- glm(canceled ~ dest, data= flights_delay, family = binomial(link = "logit"))
summary(dest.cancel.fit )
```

So from the output, all the p values are greater than 0.05, we can consider there is no significant of destination effect on cancelations.


__d) characteristics of the plane__


```{r,warning=FALSE,eval=TRUE,message=FALSE}


flights_delay <- tbl(my_db, sql("SELECT dep_delay,arr_time, year,tailnum FROM flights")
                     )%>%filter(year == 2013) %>% mutate(canceled = is.na(arr_time))
flights_delay <- flights_delay %>% left_join(planes, by = "tailnum", copy = TRUE)
tmp<- group_by(flights_delay, manufacturer) %>% summarise(meandelay = mean(dep_delay))
tmp <- as.data.frame(tmp)
fig <- ggplot(tmp, aes(x= manufacturer,y=meandelay)) + geom_bar(
  stat="identity",fill ="skyblue") + theme(axis.text.x=element_text(angle = 90))
fig
flights_delay <- as.data.frame(flights_delay)
plane.dep.fit <- lm(dep_delay ~  type + manufacturer + engines + engine, data= flights_delay)
summary(plane.dep.fit)
```

From the plot, it can be easily found, planes made by AGUSTA SPA manufacturer have the most high average departure delay time.

From the model ouput, manufacturer LAMBERT RICHARD and SIKORSKY  has significant effect on departure delays as their p values are both less than 0.05, and we can find number of engines is not significant factor, but Reciprocating type engine has a pvalue less than 0.05 which is a significant factor, At last, type is also not a significant factor.


```{r,warning=FALSE,eval=TRUE,message=FALSE}
plane.cancel.fit <- glm(canceled ~  type + manufacturer + engines + engine,
                        data= flights_delay, family = binomial(link = "logit"))
summary(plane.cancel.fit)
```

As from the output, we can see almost all of the factors are not significant factors with p values greater than 0.05, except engines has a very small p value, so it is a significant factor effect on cancelation.


In conclusion, the idea about the expectations for takeoff delays, at New York City airports in 2013, is when Temperature and Relative humidity are high and the time during some midnights in second half year, and when the destinations are for example ALB and etc. Also, if the plane is made by LAMBERT RICHARD and SIKORSKY and has a Reciprocating type engine. As for cancellations, there is a larger probability for canceled flights when there were weekends in second half year, and one more interesting condition is when the plane have many engines.


